WED-SQL (Draft)

### 2016-05-21 doc/paper/main_en.tex
paper: ch4

\par When employing the Continuous Query model in a system that is response time sensitive, one needs to keep in mind the
computational cost of this approach, which in turn depends both on the number of distinct events to be detected and on how 
often they pop up. In order to implement this model efficiently, it is critical to have a good time estimation of events
occurrence, especially in the presence of multiple distinct events, once there is an overhead cost involved in performing
a query that cannot be disregarded. In the other hand, if this time estimation is unpredictable for some or all of the
events, the suitable query must be executed as frequent as possible, otherwise threatening the system's real time response 
capability. Anyway, the Continuous Query model is not well suitable to every data event driven system, which ultimatelly
may lead to waste of computational resources, increasing the overall power consumption and possible overloading the system.

@@

\par It very important to note that a WED-worker should never rely in WED-attributes values that were not specified in
the WED-condition's predicate, once, apart from being a project error, it may end up capturing spurious conditions or refusing
to perform a legitimate one.

\subsection{Controle transacional e de concorrência}
%wed-trans, why lock ?, types of locks , why advisory(pg_row_locks)
%locks are needed for: 1-keep track of wed-trans; 2-concurrency control between wed-workers (pg_locks, select for update nowait)

Antes que uma WED-transition possa ser executada, é necessário que a respectiva tarefa esteja registrada na tabela Job\_Pool e que
o WED-worker que irá executá-la garanta exclusividade de execução. Isso é efeito requisitando uma trava ao WED-server na tarefa
especificada. Essa trava é necessária por dois motivos: permitir que o WED-server tenha controle sobre o tempo de vida da transação;
gerenciar conflitos entre transações concorrentes.

\par O modelo WED-flow prevê que cada WED-transition tenha um tempo limite de execução. Para atender a esse requisito, o WED-server
precisa identificar a transição que está sendo executada e monitorar o seu tempo de vida. Vale lembrar que a execução de uma
WED-transition está encapsulada em uma única transação. Um modo de forçar as transações a se identificarem é exigir que as mesmas
solicitem uma trava exclusiva ao WED-server. Outra possível solução seria identificar a transação semanticamente de acordo
com o novo WED-state gerado, e nesse caso teríamos dois problemas: o WED-server apenas seria capaz de identificar a transação
no momento da atualização da instância, fazendo com que as transações pudessem ficar ativas muito além de seu tempo limite para
só então serem finalizadas, potencialmente limitando o tempo máximo de execução para todas as transações; seria necessário manter 
um conjunto de possíveis estados de dados pré-definidos, limitando a capacidade do modelo em lidar com exceções. Além do mais,
o WED-SQL utiliza essas travas de execução para resolver problemas de concorrência entre WED-workers e, com isso,
é a solução adotada para a identificação das transações.

\par Cada WED-transition pode ter multiplos WED-workers idênticos associados. Isso pode ser necessário para atender a demanda 
gerada por um WED-flow com muitas instâncias. Para evitar que mais de um WED-worker execute a mesma WED-transition na mesma
instância, o mesmo deve registrar interesse em executá-la por meio da requisição ao WED-server de uma trava exclusiva na linha
correspondente da tabela Job\_Pool, que é identificada por seu WED-flow id (wid) e a WED-trigger id (tgid) que a disparou.

@@

\bibitem{FTPM10}
João E. Ferreira, Osvaldo K. Takai, Simon Malkowski e Calton Pu. 
\emph{Reducing exception handling complexity in business process modeling and implementation:
the WED-flow approach.}, Em Proceedings of the 2010 international conference on
On the move to meaningful internet systems - Volume Part I, OTM’10, páginas
150–167. Springer-Verlag, 2010.

\bibitem{ICWS12}
João Eduardo Ferreira, Kelly Rosa Braghetto, Osvaldo Kotaro Takai e
Calton Pu.\emph{ Transactional recovery support for robust exception handling in business process
services.} Em Proceedings of the 19th International Conference on Web Services (ICWS), páginas
303–310, 2012

\bibitem{SGD87}
Hector Garcia Molina, Kenneth Salem.
\emph{Sagas}, Em Proceeding of the 1987 ACM SIGMOD International Conference on Management of Data, páginas 249-259, 1987

\bibitem{CQ}
Ling Liu, Calton Pu, Wei Tang.
\emph{Continual Queries for Internet Scale Event-Driven Information Delivery},
IEEE Trans. on Knowl. and Data Eng. 11, páginas 610–628, 1999

@@

\bibitem{WT}
Marcela Ortega Garcia, Pedro Paulo de S. B. da Silva, Kelly Rosa Braghetto e João Eduardo Ferreira.
\emph{Wed-tool: uma ferramenta para o controle de execução de processos de negócio transacionais},
Proceedings of the 27th Brazillian Symposium on Databases - Demos and Applications Session, página 36, 2012

>>>>

\par When employing the Continuous Query model in a system that is response time sensitive, one needs to keep in mind the
computational cost of this approach, which in turn depends both on the number of distinct events to be detected and on how 
often they pop up. In order to implement this model efficiently, it is critical to have a good time estimation of events
occurrence, especially in the presence of multiple distinct events, once there is an overhead cost involved in performing
a query that cannot be disregarded. In the other hand, if this time estimation is unpredictable for some or all of the
events, the suitable query must be executed as frequent as possible, otherwise threatening the system's real time response 
capability. Anyway, the Continuous Query model is not well suitable to every data event driven system, which ultimately
may lead to waste of computational resources, increasing the overall power consumption and possible overloading the system.

@@

\par It is very important to note that a WED-worker should never rely in WED-attributes values that were not specified in
the WED-condition's predicate, once, apart from being a project error, it may end up capturing spurious conditions or refusing
to perform a legitimate one.

\subsection{Performing WED-transitions}
%wed-trans, why lock ?, types of locks , why advisory(pg_row_locks)
%locks are needed for: 1-keep track of wed-trans; 2-concurrency control between wed-workers (pg_locks, select for update nowait)

WED-workers can only perform WED-transitions properly recorded on the Job\_Pool table. Moreover, they must request the WED-server
an exclusive execution lock on the task to be performed, being free to proceed once this lock is granted. The WED-server rely
on this lock protocol to enforce each transaction to complete or abort within the time limits specified for each WED-transition.
It also used to solve conflicts between concurrent WED-workers.

\par One characteristic of a well designed WED-flows project is the presence of reasonable execution time limits for its WED-transitions.
Although it is up to the project designers to define these time limits, it is the role of the WED-server to enforce these restrictions.
In order to accomplish that, it must first identify a WED-transition that is about to begin, and then keep track of its running time.
It i worth to remember that each WED-transition is wrapped in a single transaction.

\par One way to force transactions to identify themselves is to require a exclusive lock when performing WED-transitions,
effectively informing the WED-server of what task in Job\_Pool is going to be worked on. Another possible solution would
be semantically identify the WED-transition based on the new WED-state that is about to be committed, but not without a 
few caveats: First, it could allow a transaction to stay alive much longer than necessary in the case when a WED-transition
would be aborted anyway due to time out. Second, the WED-server would need to know which new WED-states are valid for each
WED-transition to set, potentially harming its exception handling capability. Moreover, the locking solution was chosen 
once it can also be used to solve concurrency conflicts.

\par

@@

\bibitem{FTPM10}
João E. Ferreira, Osvaldo K. Takai, Simon Malkowski and Calton Pu. 
\emph{Reducing exception handling complexity in business process modeling and implementation:
the WED-flow approach.}, Em Proceedings of the 2010 international conference on
On the move to meaningful internet systems - Volume Part I, OTM’10, pages
150–167. Springer-Verlag, 2010.

\bibitem{ICWS12}
João Eduardo Ferreira, Kelly Rosa Braghetto, Osvaldo Kotaro Takai and
Calton Pu.\emph{ Transactional recovery support for robust exception handling in business process
services.} Em Proceedings of the 19th International Conference on Web Services (ICWS), pages
303–310, 2012

\bibitem{SGD87}
Hector Garcia Molina, Kenneth Salem.
\emph{Sagas}, Em Proceeding of the 1987 ACM SIGMOD International Conference on Management of Data, page 249-259, 1987

\bibitem{CQ}
Ling Liu, Calton Pu, Wei Tang.
\emph{Continual Queries for Internet Scale Event-Driven Information Delivery},
IEEE Trans. on Knowl. and Data Eng. 11, pages 610–628, 1999

@@

\bibitem{WT}
Marcela Ortega Garcia, Pedro Paulo de S. B. da Silva, Kelly Rosa Braghetto e João Eduardo Ferreira.
\emph{Wed-tool: uma ferramenta para o controle de execução de processos de negócio transacionais},
Proceedings of the 27th Brazillian Symposium on Databases - Demos and Applications Session, page 36, 2012
